---
title: 'BST 260: EHR'
author: Katherine Wang, Lauren Yoo, Eunice Yeh
output: html_document
---


```{r}
library(dplyr)
#https://dashboard.healthit.gov/datadashboard/documentation/ehr-products-mu-attestation-data-documentation.php
dat_mu <- read.csv("MU_REPORT_2016.csv")
#head(dat_mu)
#names(dat_mu)

# remove non-US states
dat_mu <- dat_mu %>% filter(Business_State_Territory %in% state.name |   Business_State_Territory == "District Of Columbia")

# remove zipcodes < 5
dat_mu_zip <- dat_mu %>% filter(ZIP == "" | nchar(as.character(ZIP)) == 5)

## check zipcodes
#dat_mu_ep$NPI
#which(nchar(as.character(dat_mu_ep$ZIP)) < 5)
#dat_mu %>% filter(NPI == 1164639365)
#dat_mu %>% filter(NPI == 1831180645)
#dat_mu %>% filter(NPI == 1497083901)
#dat_mu %>% filter(NPI == 1972809879)
#dat_mu %>% filter(ZIP != "" & nchar(as.character(ZIP)) < 5)
```
* removed zipcodes for US territories (zipcodes that weren't for the 50 states or DC)
* removed "zipcodes" that had length less than 5, ie zipcodes with characters and 3 digits

```{r}
library(zipcode)
data("zipcode")
dat_mu_zmerge <- dat_mu_zip %>% left_join(zipcode, by = c("ZIP" = "zip"))

#check that states match
states <- data.frame(state_name = state.name, state_abb = state.abb)

test_mismatch <- dat_mu_zmerge %>% left_join(states, by = c("state" = "state_abb")) %>% mutate(Business_State_Territory = as.character(Business_State_Territory)) %>% filter(state_name != Business_State_Territory)

# percent mismatch state
nrow(test_mismatch)/nrow(dat_mu_zip)
nrow(test_mismatch)/nrow(dat_mu_zmerge)

# check which zips are mismatching
test_mismatch %>% 
  group_by(ZIP, Business_State_Territory, state_name) %>% 
  summarize(num = n()) %>% 
  arrange(desc(num)) %>% 
  head(5)

```
about 4% mismatch in zipcode/states
top 10 mismatches:
* zipcode 51503 is in Iowa (agrees with zipcode package)/bordering Nebraska, but reported as Nebraska which is neighboring Iowa. This may be correct, so we'll keep Nebraska as reported
* zipcode 55905 is in Minnesota, but reported state Florida is not near Minnesota. Since there is likely a typo in either zip or state, we'll remove this mismatched zip all together
* zipcode 54601 is in Wisconsin/bordering Minnesota, but reported as Minnesota which is neighboring Wisconsin. We'll keep Minnesota as reported
* zipcode 33331 is in Florida, which is nowhere near Ohio. Since it is unclear which was the typo, we'll delete all these mismatches
* zipcode 55905 is in Minnesota - found it's actually Mayo Clinic (along with #2). Wisconsin and Minnesota are bordering each other, so we'll keep Wisconsin


notes: 
running this code test_mismatch %>% filter(ZIP == "55905")
notice that this zipcode is incorrectly used for multiple states, possibly meaning htere were data entry/quality issues

```{r}
#dataset removing mismatching zips
# make sure to use bus state_territory?
clean_dat <- dat_mu_zmerge %>%
  left_join(states, by = c("state" = "state_abb")) %>% 
  mutate(Business_State_Territory = as.character(Business_State_Territory)) %>%
  filter(
    (state_name == Business_State_Territory) | 
    (state_name != Business_State_Territory && ZIP == "51503" && Business_State_Territory == "Nebraska") |
    (state_name != Business_State_Territory && ZIP == "54601" && Business_State_Territory == "Minnesota") |
    (state_name != Business_State_Territory && ZIP == "55905" && Business_State_Territory == "Wisconsin")
    ) %>% 
  select(NPI, Business_State_Territory, ZIP, Specialty, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)
```


old
```{r, eval=FALSE}
# select physician columns
dat_mu_ep <- dat_mu_zip %>% filter(Provider_Type == "EP") %>% select(NPI, Business_State_Territory, ZIP, Specialty, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)

#dat_mu_ep_clean <- dat_mu_clean %>% filter(Provider_Type == "EP") %>% select(NPI, city, state, ZIP, longitude, latitude, Specialty, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)

# select hospital columns
dat_mu_hosp <- dat_mu_zip %>% filter(Provider_Type == "Hospital") %>% select(NPI, CCN, Business_State_Territory, ZIP, Hospital_Type, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)

#dat_mu_hosp_clean <- dat_mu_clean %>% filter(Provider_Type == "Hospital") %>% select(NPI, city, state, ZIP, longitude, latitude, Specialty, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)

```

##Code without Zip Package
```{r}
library(dplyr)
#https://dashboard.healthit.gov/datadashboard/documentation/ehr-products-mu-attestation-data-documentation.php
dat_mu <- read.csv("MU_REPORT_2016.csv")
#head(dat_mu)
#names(dat_mu)

# remove non-US states
dat_mu <- dat_mu %>% filter(Business_State_Territory %in% state.name |   Business_State_Territory == "District Of Columbia")

# remove zipcodes < 5
dat_mu_zip <- dat_mu %>% filter(ZIP == "" | nchar(as.character(ZIP)) == 5)

# select physician columns
dat_mu_ep <- dat_mu_zip %>% filter(Provider_Type == "EP") %>% distinct(NPI, Business_State_Territory, ZIP, Specialty, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)

# select hospital columns
dat_mu_hosp <- dat_mu_zip %>% filter(Provider_Type == "Hospital") %>% distinct(NPI, CCN, Business_State_Territory, ZIP, Hospital_Type, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)## %>% select(NPI, CCN, Business_State_Territory, ZIP, Hospital_Type, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)
```

##Exploratory
```{r}
library(tidyverse)

dat_mu_ep %>% 
  group_by(Vendor_Name) %>% 
  filter(n() > 5000) %>% 
  ggplot() + geom_bar(aes(Vendor_Name)) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

#dat_mu_ep %>% filter(n() > 5000) %>% ggplot() + geom_point(aes(x = Business_State_Territory, y = Vendor_Name, color = Vendor_Name))

dat_mu_ep %>% group_by(Vendor_Name) %>% filter(n() < 5000) %>% ggplot() + geom_bar(aes(Vendor_Name))

dat_mu_ep %>% 
  filter(agrepl("Epic", Vendor_Name)) %>% 
    ggplot() + 
    geom_bar(aes(EHR_Product_Name)) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

dat_mu_ep %>% 
  filter(agrepl("Cerner", Vendor_Name)) %>% 
    ggplot() + 
    geom_bar(aes(EHR_Product_Name)) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))



library(fiftystater)
library(RColorBrewer)
data("fifty_states")
states_map <- map_data("state")
mapplot <- dat_mu_ep %>% filter(agrepl("Cerner", Vendor_Name)) %>% group_by(Business_State_Territory) %>% summarize(count = n()) 

mapplot %>% 
  ggplot() +
  geom_map(aes(map_id = id), data = fifty_states, map = fifty_states, color="grey50", fill="grey90") +
  geom_map(aes(map_id = tolower(Business_State_Territory), fill = count), map = fifty_states, color="grey50") +
  coord_map() +
  scale_fill_gradientn(colors = brewer.pal(6, "RdYlGn")) +
  expand_limits(x=states_map$long, y=states_map$lat) +
  theme_gray() +
  theme(axis.text = element_blank(), 
        axis.line = element_blank(), 
        axis.ticks = element_blank(), 
        panel.border = element_blank(), 
        panel.grid = element_blank(), 
        axis.title = element_blank(),
        legend.position = "bottom") +
  ggtitle("Test")
```

##Save R Data
```{r}
# select physician data
dat_mu_ep <- dat_mu_zip %>% filter(Provider_Type == "EP") %>% distinct(NPI, Business_State_Territory, ZIP, Specialty, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)

# select hospital data
dat_mu_hosp <- dat_mu_zip %>% filter(Provider_Type == "Hospital") %>% distinct(NPI, CCN, Business_State_Territory, ZIP, Hospital_Type, Program_Type, Provider_Stage_Number, Vendor_Name, EHR_Product_Name, Product_Classification, Product_Setting)

saveRDS(dat_mu_ep, file = "vendor_phys.rds")
saveRDS(dat_mu_hosp, file = "vendor_hosp.rds")
```

