---
title: "Joinining hospital-affiliated physicians data and hospital demographics data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(stringr)

#phys_data <- readRDS("~/Desktop/BST260-FALL2017/BST-260-Final-Project/phys.rds")
hosp_data <- readRDS("/Users/jiminyoo/Desktop/BST260-FALL2017/BST-260-Final-Project/hosp.rds")
#vendor_hosp <- readRDS("/Users/jiminyoo/Desktop/BST260-FALL2017/BST-260-Final-Project/vendor_hosp.rds")
#vendor_phys <- readRDS("/Users/jiminyoo/Desktop/BST260-FALL2017/BST-260-Final-Project/vendor_phys.rds")
hosp_demo_data1 <- readRDS("/Users/jiminyoo/Desktop/BST260-FALL2017/BST-260-Final-Project/hosp_demo_part1.rds")
hosp_demo_data2 <- readRDS("/Users/jiminyoo/Desktop/BST260-FALL2017/BST-260-Final-Project/hosp_demo_part2.rds")
```


#Clean the hospital demographics data

```{r}
#Row-bind the hospital demographics data 1(states ak to ma) and data2(states md to wy)
hosp_demo_data <- rbind(hosp_demo_data1, hosp_demo_data2)
#1. Change the StaffedBeds,TotalDischarges, PatientDays, and Gross PatientRevenue into numeric vectors
#Incomplete: gross_patient_revenue
colnames(hosp_demo_data) <- c("Hospital_name", "City","Staffed_beds","Total_discharges", "Patient_days","Gross_patient_revenue", "State")
hosp_demo_cleaned1 = transform(hosp_demo_data, Staffed_beds = as.numeric(gsub(",", "", Staffed_beds)), 
                 Total_discharges = as.numeric(gsub(",", "", Total_discharges)),
                 Patient_days = as.numeric(gsub(",","", Patient_days)),
                 Gross_patient_revenue = as.numeric(gsub("\\$|,", "", Gross_patient_revenue)))

#if all the variables are zero, drop the hospital 
hosp_demo_cleaned2 <- hosp_demo_cleaned1[!hosp_demo_cleaned1$Gross_patient_revenue == 0, ]
setwd("~/Desktop/BST260-FALL2017/BST-260-Final-Project")
#run this later
#saveRDS(hosp_demo_cleaned2, file="hosp_demo_full.rds")
```

#Merge the cleanedhospital demographics data with the hosp_data
```{r}
#assign the most up-to-date cleaned hosp_demo data to "hosp_demo_cleaned"
hosp_demo_cleaned <- hosp_demo_cleaned2

#Use hospital name and city to merge
hosp_demo_cleaned$Hospital_name <- toupper(hosp_demo_cleaned$Hospital_name)
hosp_demo_cleaned$City_trimmed <- toupper(str_trim(hosp_demo_cleaned$City))
#hosp_demo_cleaned$City <- as.factor(hosp_demo_cleaned$City)
hosp_data$Hospital.affiliation.LBN.1 <- toupper(as.character(hosp_data$Hospital.affiliation.LBN.1))
hosp_data$City_trimmed <- toupper(str_trim(as.character(hosp_data$City)))

#Final dataset:
#joined_df_onlyHospName <- inner_join(hosp_data, hosp_demo_cleaned, by=c("Hospital.affiliation.LBN.1"="Hospital_name"))
joined_df <- inner_join(hosp_data, hosp_demo_cleaned, by=c("Hospital.affiliation.LBN.1"="Hospital_name", "City_trimmed"="City_trimmed"))

##Future discussion: what kind of join? Unique hospitals
##Number of physicians per hospital, gender prop, average grad year, most populat specialty, number of speciaties
#Remove mu_report to git push 
setwd("~/Desktop/BST260-FALL2017/BST-260-Final-Project")
saveRDS(joined_df, "Hosp_merged_data.rds")
```





















#COMPENDIUM DATA
colnames(dat_compendium)
nums <- sapply(dat_compendium, is.numeric)
numeric_cols <- colnames(dat_compendium[, nums])


hist(dat_compendium$total_mds, xlim=c(0,10000), breaks=50)
hist(dat_compendium$prim_care_mds)
hist(dat_compendium$grp_cnt)
hist(dat_compendium$hosp_cnt)
hist(dat_compendium$sys_beds)
hist(dat_compendium$sys_dsch)
hist(dat_compendium$sys_res)
hist(dat_compendium$maj_inv_owned)
hist(dat_compendium$deg_children)
hist(dat_compendium$sys_incl_majteachhosp)
hist(dat_compendium$sys_incl_vmajteachhosp)
hist(dat_compendium$sys_teachint)
hist(dat_compendium$sys_incl_highdpphosp)
hist(dat_compendium$sys_highucburden)


#################################################################################
######################Quality Measures Data#####################################
#################################################################################
#QUALITY MEASURE DATA
colnames(dat_quality)
plot(dat_quality$State)
table(dat_quality$Reporting.Program)
table(dat_quality$Measure.Name)
table(dat_quality$Rate.Definition)
table(dat_quality$Population)


##############################################################################
##########################JOIN DATASETS#######################################
##############################################################################
library(dplyr)
#MERGE BY NPI
class(hosp_data$NPI)
class(phys_data$NPI)
class(vendor_hosp$NPI)
class(vendor_phys$NPI)

#Extract unique NPIs
unique_hosp <- hosp_data[!duplicated(hosp_data$NPI),]
hosp_duplicates <- hosp_data[!duplicated(hosp_data$NPI),]
unique_phys <- phys_data[!duplicated(phys_data$NPI), ]
unique_vendor_hosp <- vendor_hosp[!duplicated(vendor_hosp$NPI), ]
unique_vendor_phys <- vendor_phys[!duplicated(vendor_phys$NPI), ]

#Combine hospital level data 
hosp_combined <-  left_join(vendor_hosp, unique_hosp, by = "NPI")
length(is.na(hosp_combined$First.Name))

#Combine physician level data
phys_combined <- left_join(vendor_phys, unique_phys, by = "NPI")
phys_inner <- inner_join(vendor_phys, phys_data, by=c("NPI"="NPI", "ZIP"="zip"))
#Extract unique NPIs from inner_joined data 
phys_dups <- phys_data[duplicated(phys_data$NPI),]
phys_dups <- phys_data[duplicated(phys_data$NPI) & duplicated(phys_data$zip), ]

nrow(phys_inner[!duplicated(phys_inner$NPI),])


#MERGE BY CCN
hosp_data$Hospital.affiliation.CCN.1 <- as.numeric(hosp_data$Hospital.affiliation.CCN.1)
inner_join(hosp_data, mu_report, by = c("Hospital.affiliation.CCN.1" = "CCN"))

##############################################################################
##########################HOSP DATA ANALYSIS###################################
##############################################################################

