---
title: "Primary and Exploratory Analysis at the Physician Level"
author: "Eunice Yeh"
output: html_document
---

Since practitioners who are affiliated with a hospital may not have a choice in using EHR or not, we will exclude these from our analysis population, which is now just the practitioners who enrolled in the Medicare Incentive Program who are not affiliated with any hospital, which come from the dataset we cleaned earlier: `phys.rds`.

```{r}
library(tidyverse)
phys <- readRDS("phys.rds")
```

## Primary Analysis: Effects of Practitioner Demographics on the Use of EHR

We need to get a unique row per practitioner in order to fit a logistic regression model on practitioner-level demographics and EHR use.
```{r}
reg <- phys %>% 
  group_by(NPI, First.Name, Last.Name, Primary.specialty, Gender, Medical.school.name, State,
           Graduation.year, Used.electronic.health.records) %>% 
  summarize(locations = n_distinct(Zip.Code)) #%>%  # the number of practices based on distinct locations
  mutate(EHRuse = case_when(Used.electronic.health.records == 'Y' ~ 1, # EHR use = 1
                            Used.electronic.health.records == '' ~ 0)) # No response to EHR use = 0
test <- phys %>% group_by(NPI) %>% summarize(nfirst = n_distinct(First.Name), 
                                             nlast = n_distinct(Last.Name), 
                                             nprim = n_distinct(Primary.specialty), 
                                             ngend = n_distinct(Gender), 
                                             nmed = n_distinct(Medical.school.name), 
                                             nstate = n_distinct(State), 
                                             ngrad = n_distinct(Graduation.year), 
                                             nehr = n_distinct(Used.electronic.health.records))
# for now do this:
reg <- phys %>% distinct(NPI, .keep_all = TRUE)
```

** describe some math to explain the theory of our model here **

FIRST look at relationships between exposure variables

```{r}
library("arulesViz")

# first need to keep certain associational variables of interest and discretize them
corr <- reg %>% ungroup() %>% filter(Medical.school.name != "OTHER") %>% 
  select(Gender, Graduation.year, Primary.specialty, State) %>% 
  mutate(gradyr = as.factor(Graduation.year),
         state = as.factor(State)
         ) %>% 
  select(-Graduation.year, -State)
View(corr)

# convert from a data frame to a transaction dataset
corrt <- as(corr, "transactions")

# create rules using the apriori
rules <- apriori(corrt, parameter=list(support=0.01, confidence=0.5))
plot(rules)
```


```{r}
subrules <- rules[quality(rules)$confidence > 0.85]
inspect(head(sort(subrules, by ="lift"),30))
plot(subrules, method="grouped", control=list(k=50))
```

This is telling us that graduation year really isn't very interesting, state is also not as interesting as medical school, primary specialty, and gender. But then we need to choose only either medical school or primary specialty beacusethey are theoretically highly correlated - you would only go to an optometry school if you want to specailize in optometry.


So this boils down to this model:

consider drawing a dag in R?
EHRuse = grad year + specialty + (gradyr x gender) + (specialty x gender)

logistic regression models:
$$
1. EHR = gradyr  \\
2. EHR = gradyr + specialty \\
3. EHR = gradyr + specialty + gender \\
4. EHR = gradyr + gradyr*gender \\
5. EHR = specialty \\
6. EHR = specialty + specialty*gender \\
7. EHR = gradyr + specialty + gradyr*gender + specialty*gender
$$


things to fix:
- calculate a new continuous variable: years since graduation (2016 - Graduation.year)
- stratify Graduation.year into tenth-tiles for the associative learning
- how to subset to primary specialty? (mental health, oncology, optometry, primary care)




```{r}
fit <- reg %>% filter(Gender !="U" & !is.na(Graduation.year) & Medical.school.name != "OTHER") %>% 
  mutate(EHRuse = case_when(Used.electronic.health.records == 'Y' ~ 1,
                            TRUE ~ 0),
         gradyr = as.factor(Graduation.year))
model1 <- glm(EHRuse ~ Gender, data = fit, family = binomial)
summary(model1)
# model2 <- glm(EHRuse ~ Graduation.year + Gender + Graduation.year:Gender, data = fit, family = binomial) # same thing as below
model2 <- glm(EHRuse ~ Graduation.year*Gender, data = fit, family = binomial)
summary(model2)
model3 <- glm(EHRuse ~ Primary.specialty, data = fit, family = binomial)
summary(model3)
```



Let's explore the relationship between med school amd primary specialty stratified by gender.

```{r}
reg %>% filter(Medical.school.name != "OTHER" & Gender !="U") %>%
  # mutate(BB_per_game = BB/G, R_per_game = R/G, HR_per_game = HR/G, HR_strata = cut(HR_per_game, breaks = quantile(HR_per_game, probs = seq(0, 1, 0.1)))) %>%  #, labels = FALSE)) %>% 
  # filter(!is.na(HR_strata)) %>% 
  group_by(Gender) %>% 
  select(Primary.specialty, Graduation.year, Gender) %>% 
  ggplot(aes(as.integer(Primary.specialty), Graduation.year)) +  
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  facet_wrap(~Gender) +
  xlab("Primary Specialty")+
  ylab("Graduation Year") +
  ggtitle("Relationship between Primary Specialty and Graduation Year stratified by Gender")

```















## Exploratory Analysis: Demographics of Practitioners Among Those Who Use EHR (and whom we have vendor information on)

Physician-level analyses will be performed on merged dataset of `phys.rds` and `vendor_phys.rds` created in then `Merge.Rmd` file, so that we are only analyzing the demographics of physicians whom we actually have specific vendor information on so the N is consistent between analyses. Will consider doing exploratory analysis on all physician data (including hospital-affiliated) if time permits.
```{r}
EHR <- readRDS("dat_merge.rds")

EHR %>% arrange(NPI) %>% select(NPI, First.Name, Last.Name,Vendor_Name, EHR_Product_Name, City, state_abb, Zip.Code, Specialty, Primary.specialty,  Used.electronic.health.records) %>% head(50)
```

Each entry in this dataset is a unique EHR product used by each practitioner at a particular location. Looks like the variable `Used.electronic.health.records` is not very accurate since there are practitioners who clearly have records of EHR product use but did not answer yes (`Y`) to having used EHR. So we will rely on the fact that we have EHR information on these practitioners to indicate whether a practitioner used EHR or not in the primary analysis (?). To be able to perform physician-level analyses, I will summarize vendor and location information for each practitioner.

```{r}
phys_EHR <- EHR %>% arrange(NPI) %>% 
  # all the characteristics considered in the group by function should already be unique at the practitioner level
  group_by(NPI, First.Name, Last.Name, Primary.specialty, Gender, Medical.school.name, state_abb,
           Graduation.year, Used.electronic.health.records) %>% 
  summarize(locations = n_distinct(Zip.Code), # number of practices (based on distinct locations) a practitioner has
            products = n_distinct(EHR_Product_Name), # number of distinct EHR products a practitioner uses
            vendors = n_distinct(Vendor_Name), # number of distinct EHR vendors a practitioner uses
            # states = n_distinct(state_abb), # number of states a practitioner uses EHR in
            EHRuse = 'Y' # Indicator for EHR use (based on the fact that we have specific EHR product info for this practitioner)
            ) %>% 
  ungroup()
head(phys_EHR, 50)

# unique(phys_EHR$states)
# looks like all practitioners practice within the same state, so will add state_abb into a grouping variable since it is a practitioner-level characteristic
```


## Exploring the Characteristics of Non-hospital-affiliated Practitioners Who Use EHR

1. Check the distributions of number of practice locations, states, products, and vendors among practitioners whom we have EHR usage information on. Counts of these distributions are the number of practitioners.

consider keeping vendor names so we can look at if the majority of the people are using only 1 vendor, is it beacuse they're one of the big vendors? and for those who have to use more than one vendor, is it because they're using smaller vendors?

2. Explore practitioner demographics: gender, graduation year, medical school attended, primary specialty.



```{r}
phys_EHR %>% group_by(Graduation.year) %>% select(products) %>% table()
```

** ADD LATER: show N's at each point of the x-axis **

```{r}
phys_EHR %>%
  ggplot(aes(factor(products), Graduation.year)) +
  geom_boxplot() + 
  facet_grid(.~Gender) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.25)) +
  xlab("Number of EHR Products") +
  ylab("Graduation Year") +
  ggtitle("Dist of Grad Year by Number of EHR Products Used by Gender")
```


3. Perform Association Rule Learning on all variables from the first two sections. --> check the validity of using this method if my variables are not independent from one another.
Here, each "transaction" is a practitioner who adapted EHR as part of the Medicare EHR Incentive Program in the U.S.

```{r}
library("arulesViz")

# first need to keep certain associational variables of interest and discretize them
assoc <- phys_EHR %>% ungroup() %>% 
  select(-NPI, -First.Name, -Last.Name, -Used.electronic.health.records, -EHRuse) %>% 
  mutate(Gradyr = as.factor(Graduation.year),
         State = as.factor(state_abb),
         locations = as.factor(locations),
         products = as.factor(products),
         vendors = as.factor(vendors)
         ) %>% 
  select(-Graduation.year, -state_abb)
View(assoc)

# convert from a data frame to a transaction dataset
assoctrans <- as(assoc, "transactions")

# create rules using the apriori
rules <- apriori(assoctrans, parameter=list(support=0.01, confidence=0.5))
plot(rules)
```

The result is a set of 4152 association rules with generally high confidence and low support (proportion of
transactions in the dataset which contain the itemset). Let's first trim this down a bit to more important rules (confidence > 0.8). Then check the top 30 rules with respect to the lift measure (a popular measure of rule strength) - the deviation of the support of the whole rule from the support
expected under independence given the supports of both sides of the rule.

```{r}
subrules <- rules[quality(rules)$confidence > 0.85]
inspect(head(sort(subrules, by ="lift"),30))
plot(subrules, method="grouped", control=list(k=50))
```
The top 18 rules are showing that the strongest associations are among practitioners who use 5 EHR products across 2 different vendors in the state of Minnesota within a single practice. The next set of top rules shows us a strong association among practitioners who use 7 EHR products across 2 different vendors in the state of California within a single practice.

We can visualize these top 18 rules using a directed network graph where the items are vertices and the directed edges are the rules in the direction of the antecedent (IF) to the consequent (THEN). (check this?)

```{r}
top18 <- head(sort(subrules, by ="lift"),18)
plot(top18, method="graph") # uses items and rules as vertices connecting them with directed edges 
plot(top18, method="graph", control=list(type="itemsets")) # uses itemsets as vertices and rules are represented by directed edges
```

Plot a parallel coordinates plot for the top 18 rules. The width of the arrows represents support and the intensity of the color represent confidence.
```{r}
plot(top18, method="paracoord", control=list(reorder=TRUE))
```


Explore these rules interactively.
```{r}
# saveAsGraph(head(sort(rules, by="lift"),100), file="physEHR.graphml")
```

